Simulating the Linux page cache in distributed applications with the WRENCH library

### 2020-12-09 paper_io_sim.tex
fixed figure 4

            Our first experiment simply runs one instance of the
            application on a single cluster node, with different input file
            sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
            to the same local disk. The goal is to validate the simulation model 
            in absence of any concurrent disk access.

            Our second experiment runs concurrent instances of the
            application on a single node, all application instances
            operating on different files of size 3~GB stored in the same
            local disk, sharing the disk bandwidth. We vary the number of
            concurrent application instances from 1 to 32 since cluster
            nodes have 32 CPU cores.

            Our third experiment occurs in the same configuration as the
            previous one, albeit reading and writing on a 400-GB NFS-mounted remote disk
            mounted from another compute node. As is commonly
            configured in HPC environments to avoid data loss, in our simulator, 
            NFS client and server read caches are enabled, but there is no cache 
            for writing on client side and server writes are configured with
            write-through cache instead of writeback. 
            Therefore, all the writes happen at disk bandwidth, but
            reads can benefit from cache hits. We simulate write-through by
            writing data directly to disk and adding that data to the server
            cache.

@@

@@

            \begin{figure*}
            \centering
            \begin{subfigure}{\linewidth}
                \centering
                   \includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
            \end{subfigure}
            \caption{Simulation errors with different input file sizes}
            \label{fig:single_error}
            \begin{subfigure}{\linewidth}
                \centering
                   \includegraphics[width=\linewidth]{result/single/figures/single_memprof.pdf}
            \end{subfigure}
            \caption{Memory profiling results with different input file sizes.
            The bars in the figures present activities in tasks (read, compute and write),
            the lines describe memory status throughout run time of the pipeline.}
            \label{fig:single_memprof}
            \begin{subfigure}{\linewidth}
                \centering
                   \includegraphics[width=\linewidth]{result/single/figures/cached_files.pdf}
            \end{subfigure}
            \label{fig:single_cache}
               \caption{Amount of file data in cache after each I/O operation. 
               \textcolor{red}{Update real results of 20GB}}
            \end{figure*}

>>>>

            Our first experiment (\textbf{Exp single-threaded}) simply runs one instance of the
            application on a single cluster node, with different input file
            sizes (20~GB, 50~GB, 75~GB, 100~GB), and with all I/Os directed
            to the same local disk. The goal is to validate the simulation model 
            in absence of any concurrent disk access.

            Our second experiment (\textbf{Exp concurrent}) runs concurrent instances of the
            application on a single node, all application instances
            operating on different files of size 3~GB stored in the same
            local disk, sharing the disk bandwidth. We vary the number of
            concurrent application instances from 1 to 32 since cluster
            nodes have 32 CPU cores.

            Our third experiment (\textbf{Exp NFS}) occurs in the same configuration as the
            previous one, albeit reading and writing on a 400-GB NFS-mounted remote disk
            mounted from another compute node. As is commonly
            configured in HPC environments to avoid data loss, in our simulator, 
            NFS client and server read caches are enabled, but there is no cache 
            for writing on client side and server writes are configured with
            write-through cache instead of writeback. 
            Therefore, all the writes happen at disk bandwidth, but
            reads can benefit from cache hits. We simulate write-through by
            writing data directly to disk and adding that data to the server
            cache.

@@

        Figure~\ref{fig:single_error} shows the relative error of simulated
        read and write times for the three successive application tasks in \textbf{Exp single-threaded}. The
        page cache simulation model drastically reduces simulation errors
        for all I/O operations except the first read as it involves
        uncached data. Errors are reduced from an average of 345\%
        in the original WRENCH to 46\% in the Python prototype and 39\% in WRENCH-cache.
        Unsurprisingly, the original WRENCH simulator significantly overestimates
        read and write times, due to its lack of page cache simulation
        model.

        Especially, the simulation time errors of \textit{Read 3} with 100 GB of input
        in the simulators with page cache are just under the error in original WRENCH.
        This longer read can be explained with the amount of cached data of files
        at the beginning of the read in reality and simulators shown in
        Fig.~\ref{fig:single_cache}.
        % In this figure, \textit{file $<$i$>$} is the input of \textit{Read $<$i$>$} and
        % output of \textit{Write $<$i-1$>$}.
        With the input size of 20 GB, the amount of cached data of each files after 
        every I/O operation is similar since the page cache can keep all these files.
%             \tristan{what does this refer to?
%             For 20 GB there is still some error in WRENCH-cache, what do you mean 
%``there is no difference in the page cache''?}.
%   Dzung: Some text was lost in a merge, I added it again.
        With the input of 100 GB, the difference between real execution and WRENCH-cache
        simulator can be seen in the amount of \textit{file 3} cache after \textit{Write 2}.
        In real execution, after being written by \textit{Write 2},
        \textit{file 3} is entirely in cache.
        Thus, in \textit{Read 3}, \textit{file 3} is read with memory bandwidth.
        However, in the WRENCH-cache simulator, only a part of \textit{file 3} is
        cached after \textit{Write 2}, which means \textit{Write 2} reads a part of
        \textit{file 3} from disk.
        The reason for the difference is in our WRENCH-cache simulator,
        with 100 GB of input, while \textit{file 3} is being written,
        the page cache is saturated, eviction is then evoked to evict data
        from page cache and make space available for \textit{file 3}.
        However, because the evicted amount is less than required,
        some \textit{file 3} data is evicted after being written, make \textit{file 3}
        partially cached.
        As a result, in \textit{Read 3}, only a part of \textit{file 3} is read from cache
        with memory bandwidth, a part of it is read from disk with disk read bandwidth,
        which is many times slower, leading to longer read time.

@@

            \begin{figure*}
            \centering
            \begin{subfigure}{\linewidth}
                \centering
                   \includegraphics[width=\linewidth]{result/single/figures/single_errors.pdf}
                   \vspace*{-0.7cm}
                   \caption{Simulation errors}
                   \vspace*{0.5cm}
                   \label{fig:single_error}
                \end{subfigure}
            \begin{subfigure}{\linewidth}
                \centering
                %    Gray shades represent task phases (read, compute and write). 
                %    Lines represent memory usage along pipeline execution time.}
                   \includegraphics[width=\linewidth]{result/single/figures/single_memprof.pdf}
                   \vspace*{-0.7cm}
                   \caption{Memory profiles}
                   \vspace*{0.5cm}
                   \label{fig:single_memprof}
            \end{subfigure}
            \begin{subfigure}{\linewidth}
                \centering
                   \includegraphics[width=1.05\linewidth]{result/single/figures/cached_files.pdf}
                   \caption{Cache contents}
                %    \textcolor{red}{Update real results of 20GB}}
                   \label{fig:single_cache}
            \end{subfigure}\\
            \caption{Single-threaded results}
            \end{figure*}
